### 2011-06-11 Text/paper.tex
Started analysis of codes results. Not very nice steady state.

The numerical  procedure of studying mass transfer is presented in the following steps:
\begin{description}
 \item[Flow field] The hydrodynamics for the bubble motion in the periodic domain is obtained
according to the work \cite{kuzmin-binary2d}. The grid resolution is taken to insure grid
independency of results.  
 \item[Bubble reference frame] Once the hydrodynamics is resolved, the mass transfer simulations
are conducted in the moving
reference frame with the bubble. Then the bubble stands still and the flow is coming around the
bubble. We impose a steady concentration of solutant on the surface of the bubble.
\item[Mass transfer] One needs not only perform a Galilean transformation to the moving reference
frame of the bubble but also to
transfer the bubble to the inlet condition. This can insure that the studied phenomenon is a liquid
slug. 
\end{description}
\begin{figure}
\includegraphics[width=\textwidth]{Figures/benchmark_new.eps}\\
\\
\\
\\
\includegraphics[width=\textwidth]{Figures/benchmark_periodic.eps}
\caption{The two-dimensional benchmarks for the hydrodynamics (top) and the mass transfer
coefficient (bottom). \label{fig:benchmark}}
\end{figure}

@@

For long slug lengths in circular capillaries the folowing correlation is derived by
\citet{bercic-mass}:
\begin{equation}
k_L a = \frac{0.111
(j_{\mathrm{gas}}+j_{\mathrm{liq}})^{1.19}}{\bigl((1-\varepsilon_{\mathrm{gas}})(L_{\mathrm{bubble}}
+L_ {\mathrm{slug}} )\bigr)^{0.57} },
\end{equation}
where $\varepsilon_{\mathrm{gas}}=\frac{U_{\mathrm{gas}}}{U_{\mathrm{bubble}}}$ is the gas holdup,
$j_{\mathrm{gas}}=U_{\mathrm{gas}} \pi R^2$ is the volumetric gas flow rate,
$j_{\mathrm{liq}}=U_{\mathrm{liq}} \pi R^2$ is the volumetric liquid flow rate. In their
experiments \citeauthor{bercic-mass} dealt with long slug lengths. Thus, they concluded that the
mass transfer from bubble caps was predominant.

@@

Assuming that the surface tension of water is $0.0728\,\mathrm{N/m}$, the density is
$1000\,\mathrm{kg/m^3}$,the hydraulic diameter of the channel to be $1.5\,\mathrm{mm}$ one can
calculate the physical velocity of the bubble.
Table \ref{table:twod:simulations} represents the number of simulations runs for the grid as
$202\times 3002$ and its nondimensional parameters. One can see that velocity is quite reasonable
as \citet{bercic-mass} declared that experiments were conducted for velocities in the ranges of
$0.01$ to $0.4\,\mathrm{m}/\mathrm{s}$ but for less capillary numbers. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|}
$Ca$&$Re$&$U_{LB}$&$\delta$&$U_{\mathrm{phys}}\mathrm{m/s}$ \\
$0.026$ &$0.449$  &$0.0014$ &$0.040$&$0.023$\\  
$0.047$ &$0.820$  &$0.0027$ &$0.058$&$0.043$\\ 
$0.080$ &$1.378$  &$0.0045$ &$0.085$&$0.073$\\
$0.065$ &$1.126$  &$0.0037$ &$0.076$&$0.059$\\
$0.222$ &$3.807$  &$0.0125$ &$0.122$&$0.202$\\  
$0.479$ &$8.222$  &$0.0271$ &$0.151$&$0.437$\\
$0.736$ &$12.617$ &$0.0416$ &$0.164$&$0.671$\\ 
$0.989$ &$16.960$ &$0.0559$ &$0.172$&$0.902$\\
\end{tabularx}\caption{Two-dimensional simulations for different
capillary numbers. \label{table:twod:simulations}}
\end{table}

\section{Lattice Boltzmann implementation}
For the lattice Boltzmann implementation we took $\omega=1.99$ which results in the following
lattice Boltzmann diffusivity parameter:
\begin{equation}
\begin{aligned}
&D=\frac{1}{3}\Bigl(\frac{1}{\omega}-\frac{1}{2}\Bigr)\\
&D=0.0008375
\end{aligned}
\end{equation}
In terms of Schmidt number (taking the liquid viscosity in former simulations as $\frac{2}{3}$) that
can result in:
\begin{equation}
\begin{aligned}
&Sc=\frac{\nu_{\mathrm{liq}}}{D}\\
&Sc=796\\
\end{aligned}
\end{equation}
Taking the water as $8.9 \times 10^{-4}\,\mathrm{Pa\cdot s}$ or kinematic viscosity as $8.9 \times
10^{-7} \,\mathrm{m^2/s}$ and the Schmidt number as $796$ that results in the following quantity:
\begin{equation}
D=1.118\times 10^{-5}\,\mathrm{cm^2/s},
\end{equation}
which is less than the diffusivity of the methane $1.84\times 10^{-5}\,\mathrm{cm^2/s}$ but is
physical. After calculation of physical parameters one can calculate characteristics for the mass
theoretical mass transfer. Table \ref{table:lengths} gives nondimensional and physical lengths of a
bubble and a slug.
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|}
Bubbl length& Slug Length\\
5.225&9.775\\
5.22&9.78\\
5.215&9.785\\
4.965&10.035\\
5.25&9.75\\
5.565&9.435\\
5.51&9.49\\
5.505&9.495\\
\end{tabularx}
\caption{The nondimensional to the height of the channel and physical lengths for the same lattice
Boltzmann simulations as in Table \ref{table:twod:simulations}. \label{table:lengths}}
\end{table}

Then the diffusion penetration length is:
\begin{equation}
\delta_{penetration}=\sqrt{D t_{film}}=\sqrt{\frac{D L_{bubble}}{U_{bubble}}}.
\end{equation}
The bubble length is $5$ of diameters. Thus, the bubble length is $7.5\,\mathrm{mm}$. Therefore:
\begin{equation}
\delta_{penetration}=\sqrt{\frac{1.84 10^{-9} \mathrm{m^2/s} 7.5 10^{-3} \mathrm{m}}{4.3 \time
10^{-2}\mathrm{m/s}}}=1.8 \times 10^{-5}
\end{equation}
With the given value of $200$ nodes to resolve the diameter for the given simulation, that results
in $2.4$ nodes, which is small but considerable number to conduct the simulation.
Foirier number which is defined as:
\begin{equation}
Fo=\frac{D t_{film}}{\delta_{film}^2}=\frac{1.84\times 10^{-9} 0.1744}{\times
(8.7\times10^{-5})^2}=0.0423
\end{equation}
As it is indicated by \citet{vanbaten-circular} the Fourier number $Fo<0.1$ stays for short
contact, the film is unsaturated. 
Thus, one can calculate the corresponding parameters:
\begin{equation}
Pe_{min}=\frac{10^{-2} 1.5 10^{-3}}{1.84  10^{-9}}=0.81 \times 10^{4}\gg 1
Sc_{min}=\frac{10^{-6}}{1.84 10^{-9}}=543,
\end{equation}
which imply the Reynolds number to be as:
\begin{equation}
Pe=Re Sc,\,Re=Pe/Sc=14.91
\end{equation}

@@

\section{LBM implementation}

@@

As far as the simulated diffusion tensor is close to $0$ the stability limits restrict the velocity
amplitude to be simulated \cite{kuzmin-stability-optimal}. However, particular choice of weights
can drastically help. Fig. \ref{stability:d2q9} shows the stability limits for D2Q9 with removed
numerical diffusion and with it for two particular choices of weights.
\begin{figure}
\includegraphics[width=\textwidth]{Figures/d2q9_stability.eps}
\caption{The non-negativity areas and stable sub-domains of the OTRT schemes ($\Lambda=\frac{1}{4}$)
are plotted for the $D2Q9$ (standard weights, left) and the $D2Q9$ (uniform weights, right) when the
numerical diffusion is removed, partially or completely. Two non-negativity areas where all the
equilibrium populations are positive have curvilinear boundaries: they are smaller when
$g^{(u)}_{xy}=0$ (yellow) and than when $g^{(u)}_{xy}=1$ (green+yellow). The non-negativity
condition is required by BGK collision operator. The stable sub-domains for $g^{(u)} = 1,
g^{(u)}_{xy} = 0$ are small (red+green+yellow) triangles. When the numerical diffusion is removed
the stable area is bounded by magenta and the (blue) domain above. The proper choice of weights can
allow to obtain  total triangle $0 \leq u_x^2+u_y^2 \leq 1−c_e$ (the
uniform weights only).
\label{stability:d2q9}}
\end{figure}

@@

\subsection{LBM derivation}

>>>>

The numerical  procedure of studying mass transfer is presented in the following steps:
\begin{description}
 \item[Flow field] The hydrodynamics for the bubble motion in the periodic domain is obtained
according to the work \cite{kuzmin-binary2d}. The grid resolution is taken to insure grid
independency of results.  
 \item[Bubble reference frame] Once the hydrodynamics is resolved, the mass transfer simulations
are conducted in the moving
reference frame with the bubble. Then the bubble stands still and the flow is coming around the
bubble. We impose a steady concentration of solutant on the surface of the bubble.
\item[Mass transfer] One needs not only perform a Galilean transformation to the moving reference
frame of the bubble but also to
transfer the bubble to the inlet condition. This can insure that the studied phenomenon is a liquid
slug. 
\end{description}
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_new.eps}\\
\\
\\
\\
\includegraphics[width=\textwidth]{Figures/benchmark_periodic.eps}
\caption{The two-dimensional benchmarks for the hydrodynamics (top) and the mass transfer
coefficient (bottom). \label{fig:benchmark}}
\end{figure}

@@

For long slug lengths in circular capillaries the folowing correlation is derived by
\citet{bercic-mass}:
\begin{equation}
k_L a = \frac{0.111
(U_{\mathrm{gas}}+U_{\mathrm{liq}})^{1.19}}{\bigl((1-\varepsilon_{\mathrm{gas}})(L_{\mathrm{bubble}}
+L_ {\mathrm{slug}} )\bigr)^{0.57} },
\end{equation}
where $\varepsilon_{\mathrm{gas}}=\frac{U_{\mathrm{gas}}}{U_{\mathrm{bubble}}}$ is the gas holdup,
$U_{\mathrm{gas}}$ is the superficial gas velocity, and $U_{\mathrm{liq}}$ is the superficial
liquid velocity. The superficial liquid velocity is calculated with the steady state condition in
the middle of the slug as:
\begin{equation}
U_{\mathrm{liq}}=\frac{\oint U \mathrm{d}A}{A}.
\end{equation}
The superficial liquid velocity can be calculated through the calculated bubble volume and gas
holdup:
\begin{equation}
\begin{aligned}
&\varepsilon_{\mathrm{gas}}=\frac{\text{bubble area}}{15 H^2}\\
&U_{\mathrm{gas}}=\varepsilon_{\mathrm{gas}} U_{\mathrm{bubble}}
\end{aligned}
\end{equation}

In their experiments \citeauthor{bercic-mass} dealt with long slug lengths.
Thus, they concluded that the mass transfer from bubble caps was predominant.

@@

Assuming that the surface tension of water is $0.0728\,\mathrm{N/m}$, the density is
$1000\,\mathrm{kg/m^3}$,the hydraulic diameter of the channel to be $1.5\,\mathrm{mm}$ one can
calculate the physical velocity of the bubble.
Table \ref{table:twod:simulations} represents the number of simulations runs for the grid as
$202\times 3002$ and its nondimensional parameters. One can see that velocity is quite reasonable
as \citet{bercic-mass} declared that experiments were conducted for velocities in the ranges of
$0.01$ to $0.4\,\mathrm{m}/\mathrm{s}$ but for less capillary numbers. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
$Ca$    &$Re$     &$U_{LB}$ &$\delta$&$\varepsilon_{\mathrm{gas}}$
&$U_{\mathrm{phys}}\mathrm{m/s}$\\
$0.026$ &$0.449$  &$0.0014$ &$0.040$ &$0.306$ &$0.023$   \\ 
$0.047$ &$0.820$  &$0.0027$ &$0.058$ &$0.293$ &$0.043$   \\ 
$0.080$ &$1.378$  &$0.0045$ &$0.085$ &$0.280$ &$0.073$ \\
$0.065$ &$1.126$  &$0.0037$ &$0.076$ &$0.266$ &$0.059$      \\
$0.222$ &$3.807$  &$0.0125$ &$0.122$ &$0.253$ &$0.202$  \\
$0.479$ &$8.222$  &$0.0271$ &$0.151$ &$0.249$ &$0.437$  \\
$0.736$ &$12.617$ &$0.0416$ &$0.164$ &$0.236$ &$0.671$  \\ 
$0.989$ &$16.960$ &$0.0559$ &$0.172$ &$0.230$ &$0.902$  \\
\end{tabularx}
\caption{Two-dimensional simulations for different
capillary numbers. \label{table:twod:simulations}}
\end{table}

\section{Lattice Boltzmann implementation}
For the lattice Boltzmann implementation we took $\omega=1.99$ which results in the following
lattice Boltzmann diffusivity parameter:
\begin{equation}
\begin{aligned}
&D=\frac{1}{3}\Bigl(\frac{1}{\omega}-\frac{1}{2}\Bigr)\\
&D=0.0008375
\end{aligned}
\end{equation}
In terms of Schmidt number (taking the liquid viscosity in former simulations as $\frac{2}{3}$) that
can result in:
\begin{equation}
\begin{aligned}
&Sc=\frac{\nu_{\mathrm{liq}}}{D}\\
&Sc=796\\
\end{aligned}
\end{equation}
Taking the water as $8.9 \times 10^{-4}\,\mathrm{Pa\cdot s}$ or kinematic viscosity as $8.9 \times
10^{-7} \,\mathrm{m^2/s}$ and the Schmidt number as $796$ that results in the following quantity:
\begin{equation}
D=1.118\times 10^{-5}\,\mathrm{cm^2/s},
\end{equation}
which is less than the diffusivity of the methane $1.84\times 10^{-5}\,\mathrm{cm^2/s}$ but is
physical. After calculation of physical parameters one can calculate characteristics for the mass
theoretical mass transfer. Table \ref{table:lengths} gives nondimensional and physical lengths of a
bubble and a slug.
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$Ca$&Bubble length& Slug Length&
$U_{\mathrm{gas}}\,\mathrm{m/s}$&$U_{\mathrm{liq}}\,\mathrm{m/s}$&$a,\,\mathrm{m^{-1}}$\\
\hline
$0.026$&$5.225$&$9.775$ &$0.0070$ &$0.0209$&$2.5538$\\
$0.047$&$5.22$ &$9.78$  &$0.0126$ &$0.0379$&$2.5431$\\
$0.080$&$5.215$&$9.785$ &$0.0204$ &$0.0615$&$2.5326$\\
$0.065$&$4.965$&$10.035$&$0.0157$ &$0.0504$&$2.4142$\\
$0.222$&$5.25$ &$9.75$  &$0.0511$ &$0.1560$&$2.5253$\\
$0.479$&$5.565$&$9.435$ &$0.1089$ &$0.3155$&$2.6518$\\
$0.736$&$5.51$ &$9.49$  &$0.1584$ &$0.4677$&$2.6269$\\
$0.989$&$5.505$&$9.495$ &$0.2074$ &$0.6190$&$2.6340$\\
\hline
\end{tabularx}

\caption{The nondimensional (scaled to the height of the channel) lengths for the same
lattice Boltzmann simulations as in Table \ref{table:twod:simulations}. As well superficial
gas-liquid velocities are presented.\label{table:lengths}}
\end{table}

The literature correlations of mass transfer dependence on the bubble velocities are presented in
Fig. \ref{fig:mass:transfer:theoretical}. Note, that the correlations presented in
Fig.\ref{fig:mass:transfer:theoretical} are used primarily for circular capillaries
(\citeauthor{vanbaten-circular,bercic-mass}) and for square capillaries (\citeauthor{yue-mass}). We
adapt these results for flow between parallel plats by using hydraulic
diameter \cite{bercic-mass} which can be calculated as:
\begin{equation}
d_h = \lim_{W\to\infty}\frac{4A}{P}=\lim_{W\to\infty}\frac{4 W\,H}{2(W+H)}=2 H.
\end{equation}
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/theoretical_correlations.eps}
\caption{The literature correlations by various authors. One
can see the large deviation among them.\label{fig:mass:transfer:theoretical}}
\end{figure}

The volumetric mass transfer coefficient is calculated according to the work of
\citet{vanbaten-circular}. The concentration flux is calculated as the difference between overall
average concentration taken in the whole domain ($C_{overall}=\int_{V} C \mathrm{d}V /V$) at time
$t_1$ and time $t_2$ divided on the time difference $t_2-t_1$. Then the volumetric mass transfer
coefficient is calculated as:
\begin{equation}
\label{main:simulation:equation}
k_L a=\frac{\text{Flux}}{C_s-C_{outlet}} \frac{\text{bubble surface area}}{\text{unit cell volume}},
\end{equation}
where $C_{outlet}=\int{C U_{outlet}\mathrm{d}A}/\int{U_{outlet}\mathrm{d}A}$. To properly do
non-dimensioanalization one needs to insure the time conversion factor when the flux is calculated
inside the lattice Botlzmann system. The time conversion can be easily obtained:
\begin{equation}
\begin{aligned}
U_{LB}=U_{phys}\frac{\Delta t}{\Delta x}\\
\Delta t=\frac{U_{LB}}{U_{phys}}\Delta x\\
\Delta t=1.5\,\mathrm{mm} \frac{0.0559}{0.902\,\mathrm{m/s}}=9.3\times 10^{-5}\,\mathrm{s} 
\end{aligned}
\end{equation}

The bubble surface length is the perimeter of the bubble in two-dimensional space and it is
calculated with the extraction of the bubble countour and representing it with the polygon. After
that the calculated perimeter is the effective summation of line lengths and ca be represented as:
\begin{equation}
P=\sum_i{\sqrt{(x2_i-x1_i)^2+(y2_i-y1_i)^2}},
\end{equation}
where the index $i$ represents the polyline consisting of two points $(x1,y1)$ and $(x2,y2)$.

The corresponding volumetric mass transfer coefficients depending on the velocity are indicated in
Table \ref{table:lengths}.

\section{Simulation results}

\subsection{Bubble velocity dependence}
After calculation of the results and using Eq.\ref{main:simulation:equation} One can see the mass
transfer coefficients indicated in Fig. \ref{}.

\section{To revise:}
Then the diffusion penetration length is:
\begin{equation}
\delta_{penetration}=\sqrt{D t_{film}}=\sqrt{\frac{D L_{bubble}}{U_{bubble}}}.
\end{equation}
The bubble length is $5$ of diameters. Thus, the bubble length is $7.5\,\mathrm{mm}$. Therefore:
\begin{equation}
\delta_{penetration}=\sqrt{\frac{1.84 10^{-9} \mathrm{m^2/s} 7.5 10^{-3} \mathrm{m}}{4.3 \time
10^{-2}\mathrm{m/s}}}=1.8 \times 10^{-5}
\end{equation}
With the given value of $200$ nodes to resolve the diameter for the given simulation, that results
in $2.4$ nodes, which is small but considerable number to conduct the simulation.
Foirier number which is defined as:
\begin{equation}
Fo=\frac{D t_{film}}{\delta_{film}^2}=\frac{1.84\times 10^{-9} 0.1744}{\times
(8.7\times10^{-5})^2}=0.0423
\end{equation}
As it is indicated by \citet{vanbaten-circular} the Fourier number $Fo<0.1$ stays for short
contact, the film is unsaturated. 
Thus, one can calculate the corresponding parameters:
\begin{equation}
Pe_{min}=\frac{10^{-2} 1.5 10^{-3}}{1.84  10^{-9}}=0.81 \times 10^{4}\gg 1
Sc_{min}=\frac{10^{-6}}{1.84 10^{-9}}=543,
\end{equation}
which imply the Reynolds number to be as:
\begin{equation}
Pe=Re Sc,\,Re=Pe/Sc=14.91
\end{equation}

@@

@@

As far as the simulated diffusion tensor is close to $0$ the stability limits restrict the velocity
amplitude to be simulated \cite{kuzmin-stability-optimal}. However, particular choice of weights
can drastically help. Fig. \ref{stability:d2q9} shows the stability limits for D2Q9 with removed
numerical diffusion and with it for two particular choices of weights.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/d2q9_stability.eps}
\caption{The non-negativity areas and stable sub-domains of the OTRT schemes ($\Lambda=\frac{1}{4}$)
are plotted for the $D2Q9$ (standard weights, left) and the $D2Q9$ (uniform weights, right) when the
numerical diffusion is removed, partially or completely. Two non-negativity areas where all the
equilibrium populations are positive have curvilinear boundaries: they are smaller when
$g^{(u)}_{xy}=0$ (yellow) and than when $g^{(u)}_{xy}=1$ (green+yellow). The non-negativity
condition is required by BGK collision operator. The stable sub-domains for $g^{(u)} = 1,
g^{(u)}_{xy} = 0$ are small (red+green+yellow) triangles. When the numerical diffusion is removed
the stable area is bounded by magenta and the (blue) domain above. The proper choice of weights can
allow to obtain  total triangle $0 \leq u_x^2+u_y^2 \leq 1−c_e$ (the
uniform weights only).
\label{stability:d2q9}}
\end{figure}

@@
